Query Optimization for Dynamic Imputation

### 2017-03-01 paper/experiments.tex
Update query numbers

\subsection{Data sets}\label{subsec:datasets}
We collected three data sets for our experiments.
For all data sets, we selected a subset of the original attributes.
We also transformed all data values to an integer representation by enumerating strings and transforming floating-point values into an appropriate range.

@@

Our first set of queries is on the CDC data (\Cref{fig:queries-cdc}).
\Cref{q1} calculates calculate the average cuff size for users based on their income data.
\Cref{q2} compares creatine levels for individuals with low, medium, and high incomes and above a certain weight.
\Cref{q3} extracts the average blood lead levels for children under the age of 6 years.
\Cref{q4} calculates the average systolic blood pressure, by gender, for subjects with a body mass index indicating obesity.

Our second set of queries is on the freeCodeCamp data (\Cref{fig:queries-fcc}).
\Cref{q5} calculates the average income for survey participants, based on their bootcamp attendance.
\Cref{q6} estimates the average commute time of women from the United States who participated.
\Cref{q7} calculates the average amount of student debt based on school degree for survey participants who have student debt.
\Cref{q8} joins the freeCodeCamp data with a reference table provided by the World Bank which summarizes GDP per-capita across various countries~\cite{worldbank-data}.
The query calculates the average GDP per-capita of countries with and without bootcamp participants.

\begin{table*}
\todo{update this with final queries}
  \centering
  \begin{subtable}{\linewidth}
    \input{queries-cdc}
    \caption{Queries on CDC data}\label{fig:queries-cdc}
  \end{subtable}
  \par\medskip
  \begin{subtable}{\linewidth}
    \input{queries-fcc}
    \caption{Queries on freeCodeCamp data}\label{fig:queries-fcc}
  \end{subtable}
  \par\medskip
  \caption{Queries used in our experiments.}\label{fig:queries}
\end{table*}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/running_times_combined_bar_13.png}
\caption{Query runtimes with imputation on base tables, \ProjectName{}
    optimizing for quality ($\alpha=0$), and \ProjectName{} optimizing for
    performance ($\alpha=1$). Quality-optimized, on-the-fly imputation provides an order of
    magnitude speedup over imputing on base tables. Efficiency-optimized on-the-fly
    imputation can provide another order-of-magnitude speedup at the expense of quality.}

@@

\Cref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE)~\cite{Makridakis2000451} for
\ProjectName{}'s query results when compared to running imputation on the base tables and
executing the query on this cleaned base data~\cite{tofallis2015better}.  Each query was
separately run 200 times in both settings and then results from the two approaches were
paired up and compared tuple-wise.  We average tuple-wise absolute percentage deviations
within each iteration of a query, and we report this value averaged over all iterations.  We
can see that optimizing for quality indeed reduces the SMAPE of query results.  In general, the SMAPE relative 
to the base-imputation approach are low in all cases --- between
\lowsmapealphazero{} and \highsmapealphaoneexacs{} percent ---, indicating that on-the-fly
imputation produces similar results to imputation at the base tables.

We also calculate SMAPE over the number of tuples used to produce the results
in each configuration with respect to the number of tuples satisfying the same query when
the base tables have been imputed. In many cases where the SMAPE reduction between $\alpha = 1$ and
$\alpha = 0$ is small, for example \Cref{q6}, we highlight that the count of tuples can be significantly different between
the two scenarios. So while the aggregate used might not have been significantly affected by the different population,
other aggregates may well be. This highlights a challenge for a user handling data imputation traditionally: it is unclear if
the missing data will have a large or small negative impact on their analysis until they have paid the cost of
running it. By using \ProjectName{} this cost can be lowered significantly.

\begin{table}
\centering
\input{tables/perf_summary_13.tex}
\caption{Symmetric-Mean-Absolute-Percentage-Error for queries run under different $\alpha$
    parameterizations relative to results when imputing on the base table. Queries optimized
    for quality ($\alpha=0$) generally achieve lower error than queries optimized for
    efficiency ($\alpha=1$). Count shares show the number of results returned from each
    query as a percentage of the number of results returned when imputing on the base table.
    A lower count share reflects more potential for errors.}
\label{table:smape}
\end{table}

>>>>

\subsection{Data sets}\label{subsec:datasets}
We collected three data sets for our experiments.
For all data sets, we selected a subset of the original attributes.
We also transformed all data values to an integer representation by enumerating strings and transforming floating point values into an appropriate range.

@@

Our first set of queries is on the CDC data (\Cref{fig:queries-cdc}).
\Cref{q1} calculates calculate the average cuff size for users based on their income data.
\Cref{q2} compares creatine levels for individuals with low, medium, and high incomes and above a certain weight.
\Cref{q3} extracts the average blood lead levels for children under 6 years of age.
\Cref{q4} calculates the average systolic blood pressure, by gender, for subjects with a body mass index indicating obesity. 
\Cref{q5} calculates the average waist circumference for subjects above a certain height and
weight.

Our second set of queries is on the freeCodeCamp data (\Cref{fig:queries-fcc}).
\Cref{q6} calculates the average income for higher-income survey participants, grouped by their bootcamp attendance.
\Cref{q7} estimates the average commute time of women from the United States who participated.
\Cref{q8} calculates the average amount of student debt based on school degree for survey participants who have student debt.
\Cref{q9} joins the freeCodeCamp data with a reference table provided by the World Bank which summarizes GDP per-capita across various countries~\cite{worldbank-data}.
The query calculates the average GDP per-capita of countries with and without bootcamp
participants over 18 years of age.

\begin{table*}
  \centering
  \begin{subtable}{\linewidth}
    \input{queries-cdc}
    \caption{Queries on CDC data}\label{fig:queries-cdc}
  \end{subtable}
  \par\medskip
  \begin{subtable}{\linewidth}
    \input{queries-fcc}
    \caption{Queries on freeCodeCamp data}\label{fig:queries-fcc}
  \end{subtable}
  \par\medskip
  \caption{Queries used in our experiments.}\label{fig:queries}
\end{table*}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/running_times_combined_bar.png}
\caption{Query runtimes with imputation on base tables, \ProjectName{}
    optimizing for quality ($\alpha=0$), and \ProjectName{} optimizing for
    performance ($\alpha=1$). Quality-optimized, on-the-fly imputation provides an order of
    magnitude speedup over imputing on base tables. Efficiency-optimized on-the-fly
    imputation can provide another order-of-magnitude speedup at the expense of quality.}

@@

\Cref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE)~\cite{Makridakis2000451} for
\ProjectName{}'s query results when compared to running imputation on the base tables and
executing the query on this cleaned base data \cite{tofallis2015better}.  Each query was
separately run 200 times in both settings and then results from the two approaches were
paired up and compared tuple-wise.  We average tuple-wise absolute percentage deviations
within each iteration of a query, and we report this value averaged over all iterations.  We
can see that optimizing for quality indeed reduces the SMAPE of query results.  In general, the SMAPE relative 
to the base-imputation approach are low in all cases --- between
\lowsmapealphazero{} and \highsmapealphaoneexacs{} percent ---, indicating that on-the-fly
imputation produces similar results to imputation at the base tables.

We also calculate SMAPE over the number of tuples used to produce the results
in each configuration with respect to the number of tuples satisfying the same query when
the base tables have been imputed. In many cases where the SMAPE reduction between $\alpha = 1$ and
$\alpha = 0$ is small, for example \Cref{q6}\todo{confirm}, we highlight that the count of tuples can be significantly different between
the two scenarios. So while the aggregate used might not have been significantly affected by the different population,
other aggregates may well be. This highlights a challenge for a user handling data imputation traditionally: it is unclear if
the missing data will have a large or small negative impact on their analysis until they have paid the cost of
running it. By using \ProjectName{} this cost can be lowered significantly.

\begin{table}
\centering
\input{tables/perf_summary.tex}
\caption{Symmetric-Mean-Absolute-Percentage-Error for queries run under different $\alpha$
    parameterizations relative to results when imputing on the base table. Queries optimized
    for quality ($\alpha=0$) generally achieve lower error than queries optimized for
    efficiency ($\alpha=1$). Count shares show the number of results returned from each
    query as a percentage of the number of results returned when imputing on the base table.
    A lower count share reflects more potential for errors.}
\label{table:smape}
\end{table}
